---
title: "STAT 433 HW 1"
output:
  github_document
date: '2022-09-20'
---

Package Loading:
```{r, message=FALSE}
library(dplyr)
library(nycflights13)
library(ggplot2)
library(tidyverse)
```

### Question 1
**How many flights have a missing dep_time? **

```{r}
misfl = flights %>%
  filter(is.na(dep_time))

nrow(misfl)
```

**What other variables are missing?**
```{r}
misfl %>%
  select(dep_time, dep_delay, arr_time, air_time)
```

The missing columns are departure delay, arrival time and arrival delay.

**What might these rows represent?**

These values could represent that the flights were cancelled.

### Question 2
**Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.**

```{r}
flights %>%
  mutate(sched_dep_time_mins = (hour*60) + (minute)) %>%
  mutate(dep_time_mins = sched_dep_time_mins + dep_delay) %>%
  select(sched_dep_time_mins, dep_time_mins)
```

### Question 3
**Look at the number of canceled flights per day. Is there a pattern? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()**

```{r, message=FALSE}
flights %>%
  mutate(cancelled = (is.na(arr_delay) & is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(cancelled_flights = sum(cancelled), total_flights = n()) %>%
  ggplot() +
  geom_point(aes(y = cancelled_flights, x = total_flights))
```
From the plot, it suggests that as the total number of flights increase, the amound of delayed flights likewise increase.

**Is the proportion of canceled flights related to the average delay? **

```{r, message = FALSE}
flights %>%
  mutate(cancelled = (is.na(arr_delay) & is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(mean_cancelled = mean(cancelled),
            avg_dep_deplay = mean(dep_delay, na.rm = T)) %>%
  ggplot()+
  geom_point(aes(avg_dep_deplay, mean_cancelled))
```
From the plot, it seems like with longer departure delays, the number of cancelled flights on average increases.